Effect of quality-adjusted production function inputs

### 2015-05-20 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Now including dynamic SR graph.

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
commonVars <- c("Year", "Country", "Source")
historicalData$PT <- filter(IST, Country == "PT") 
historicalData$PT <- historicalData$PT[ , unique(c(commonVars,
                                                   all.vars(formulas$PT$unadjusted), 
                                                   all.vars(formulas$PT$adjusted)))]
historicalData$UK <- filter(Leeds, Country == "UK")
historicalData$UK <- historicalData$UK[ , unique(c(commonVars,
                                                   all.vars(formulas$UK$unadjusted), 
                                                   all.vars(formulas$UK$adjusted)))]
# allHistoricalData <- do.call(plyr::rbind.fill, historicalData)
@

@@

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,2, 3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  flavors <- c("unadjusted", "adjusted")
  nests <- NULL
  nests$kle <- c(1,2,3); nests$lek <- c(2,3,1); nests$ekl <- c(3,1,2)
  energies <- c(FALSE, TRUE)

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, data = historicalData[[country]], nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  coeffs <- naturalCoef(mod)
  res <- rbind(list(Country=Country, flavor=flavor, energy=energy, model=model, nest=nest))
  res <- cbind(res, coeffs)
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
@

@@

<<Dynamic Solow Residual>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
lag <- 1       # Desired lag for calculations. Set to n for differences across n years.
pad <- "tail"  # Desired padding for NA values

fd <- leaf_apply(formulas, class = "formula", f = function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]

  denom <- summands(form, n=length(summands(form))) # time variable
  nums <- c(summands(form, n=1, right = FALSE), summands(form, n=1:3)) # y, k, l, and e
  trimmedData <- historicalData[[Country]][c(denom, nums)] # columns are: iYear, y, k, l, and e
  df <- list()
  df <- historicalData[[Country]][ , c("Year", "Country")]
  df$dydt <- ediff(trimmedData[[2]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dkdt <- ediff(trimmedData[[3]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dldt <- ediff(trimmedData[[4]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
  df$dedt <- ediff(trimmedData[[5]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)

  df$flavor <- flavor
  return(df)
}) %>% do.call(rbind, .)

# origCDKLCoeffs <- filter(AllCoeffs, method=="orig", model=="cd")
# origCDKLCoeffs <- origCDKLCoeffs %>% select(method, country, Source, nestStr, energy, lambda, alpha, beta, gamma)
# # Change name of Source REXS to REXS1960 and relevel to conform to information in fdlambda.
# origCDKLCoeffs$Source <- as.character(origCDKLCoeffs$Source)
# origCDKLCoeffs$Source <- replace(origCDKLCoeffs$Source, which(origCDKLCoeffs$Source=="REXS"), "REXS1960")
# origCDKLCoeffs$Source <- as.factor(origCDKLCoeffs$Source)
# origCDKLCoeffs$Source <- relevelFactor(origCDKLCoeffs$Source, levs = c("REXS1960", "IST", "Leeds"))
# 
# # Fill new columns in the fd data frame with 
# # lambda, alpha, beta, and gamma values from the parameter estimation process.
# # lambda, alpha, beta, and gamma values for the orig fits are in the origCDKLCoeffs data frame.
# # Using the hint at 
# # http://stackoverflow.com/questions/15059076/
# # r-how-to-call-apply-like-function-on-each-row-of-dataframe-with-multiple-argum
# labg <- apply(fd[ , c('Country', 'Source', 'energyType')], 1, function(x){
#     row <- as.data.frame(filter(origCDKLCoeffs, 
#                                 country==x["Country"], Source==x["Source"], energy==x["energyType"]))
#     abg <- subset(row, select=c(lambda, alpha, beta, gamma))
#     if (nrow(abg) < 1){ abg[1, ] <- NA }
#     return(abg)
#   })
# cols <- do.call(rbind, labg)
# fd <- cbind(fd, cols)
# 
# # Now calculate lambda by finite difference for every year to make a time series
# fdlambda <- mutate(fd, lambda_diff = dydt/iGDP - (alpha/iK*dkdt + beta/iL*dldt + gamma/iE*dedt))
# 
# # Calculate the fraction of GDP growth in each year that is explained by the solow residual
# fdlambda <- mutate(fdlambda, solowFrac = lambda_diff / (dydt/iGDP))
# 
# # Calculate average values for lambda and solowFrac
# avgs <- fdlambda %>% 
#   group_by(Source, Country, energyType) %>% 
#   summarise(
#     avg.lambda_diff = mean(lambda_diff, na.rm=TRUE),
#     avg.solowFrac = mean(solowFrac, na.rm=TRUE)
#     )

@

>>>>

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
commonVars <- c("Year", "Country", "Source")
historicalData$PT <- filter(IST, Country == "PT") 
historicalData$PT <- historicalData$PT[ , unique(c(commonVars,
                                                   all.vars(formulas$PT$unadjusted), 
                                                   all.vars(formulas$PT$adjusted)))]
historicalData$UK <- filter(Leeds, Country == "UK")
historicalData$UK <- historicalData$UK[ , unique(c(commonVars,
                                                   all.vars(formulas$UK$unadjusted), 
                                                   all.vars(formulas$UK$adjusted)))]
@

@@

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot of pre-econometric data
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,2, 3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Models>>=
fileName <- "ESEE2015Models.Rdata"
if (!rerunModels){
  # load from the file on disk
  load(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  flavors <- c("unadjusted", "adjusted")
  nests <- NULL
  nests$kle <- c(1,2,3); nests$lek <- c(2,3,1); nests$ekl <- c(3,1,2)

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- 
        lapply(nests, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  save(models, file=fileName)
}
# Put all the coefficients into a data frame
AllCoeffs <- leaf_apply(models, class = c("CDEmodel", "cesModel", "LINEXmodel"), f = function(mod, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  Country <- id_list[[1]]
  flavor <- id_list[[2]]
  energy <- id_list[[3]]
  model <- id_list[[4]]
  nest <- if(length(id_list) > 4) id_list[[5]] else NA
  coeffs <- naturalCoef(mod)
  res <- rbind(list(Country=Country, flavor=flavor, energy=energy, model=model, nest=nest))
  res <- cbind(res, coeffs)
  return(res)
}) %>% do.call(plyr::rbind.fill, .)
# Relevel factors
AllCoeffs$Country <- relevelFactor(as.factor(unlist(AllCoeffs$Country)), c("PT", "UK"))
AllCoeffs$flavor <- relevelFactor(as.factor(unlist(AllCoeffs$flavor)), c("unadjusted", "adjusted"))
AllCoeffs$energy <- relevelFactor(as.factor(unlist(AllCoeffs$energy)), c("noE", "withE"))
AllCoeffs$model <- relevelFactor(as.factor(unlist(AllCoeffs$model)), c("CD", "CES"))
AllCoeffs$nest <- relevelFactor(as.factor(unlist(AllCoeffs$nest)), c("kl", "kle", "lek", "ekl"))
@

@@

<<Dynamic Solow Residuals>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
lag <- 2       # Desired lag for calculations. Set to n for differences across n years.
pad <- "tail"  # Desired padding for NA values

flavors <- c("unadjusted", "adjusted")
# model.types <- c("CD", "CES")
model.types <- c("CD")
nests <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))

dframes <- list()
for (Coun in countries){
  for (flav in flavors){
    forms <- list(noE = keep_summands(formulas[[Coun]][[flav]], -3), withE = formulas[[Coun]][[flav]])
    for (form in forms){
      denom <- summands(form, n=length(summands(form))) # time variable
      nums <- all.vars(form) [1:(length(all.vars(form))-1)] # y, k, l, and e (if present)
      trimmedData <- historicalData[[Coun]][c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)
      df <- historicalData[[Coun]][ , c(nums, "Year")]
      # Calculate names for the columns of finite difference derivatives
      derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], "/", "dt")}) %>% unlist
      derivCols <- lapply(1:length(nums), function(i){
        # Calculate finite difference derivative columns
        ediff(trimmedData[[i+1]], pad=pad, lag=lag) / ediff(trimmedData[[1]], pad=pad, lag=lag)
      }) %>% do.call(cbind, .) %>% as.data.frame
      names(derivCols) <- derivNames
      df <- cbind(derivCols, df)
      e <- if (length(summands(form))==4) "withE" else "noE"
      for (m in model.types){
        origCoeffs <- filter(AllCoeffs, model==m & energy==e & Country==Coun & flavor==flav)
        df <- cbind(df, origCoeffs)
        if (m == "CD"){
          lambda <- lapply(1:length(derivNames), function(i){
            # derivative term divided by its numerator.
            # derivative term is in the ith column of df
            # numerator is in i + length(derivNames) th column of df
            out <- df[[i]] / df[[i+length(derivNames)]]
            if (i > 1){
              # Need to make it negative and multiply by the corresponding alpha term.
              # Typical setup is that columns 2 is capital, and we need alpha_1 for capital.
              # So, subtract 1 from i.
              out <- - out * df[[paste0("alpha_", i-1)]]
            }
            return(out)
          }) %>% as.data.frame %>% rowSums
          df$lambda_fd <- lambda
          avg.lambda_fd <- mean(df$lambda_fd, na.rm = TRUE)
          # Fraction of growth explained by Solow Residual
          df$solowFrac <- df$lambda_fd / (df[[1]] / df[[length(derivNames) + 1]])
          avg.solowFrac <- mean(df$solowFrac, na.rm = TRUE)
        } else if (m == "CES"){

        } else {
          stop(paste("Unknown model", model))
        }
        df$flavor <- if (flav == "adjusted") "Quality-adjusted" else "Unadjusted"
        df$flavor <- relevelFactor(as.factor(df$flavor), c("Unadjusted", "Quality-adjusted"))
        df$energy <- if (e == "noE") "Without energy" else "With energy"
        df$energy <- relevelFactor(as.factor(df$energy), c("Without energy", "With energy"))
        dframes[[length(dframes)+1]] <- df
      }
    }
  }      
}
solowResiduals <- do.call(plyr::rbind.fill, dframes)
@

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow Residual~($\lambda$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with timespan for finite differences set to 
\Sexpr{lag}~\Sexpr{if (lag == 1) "year" else "years"}.

<<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model.">>=
data <- filter(solowResiduals, model=="CD")
ggplot(data=data) +
  geom_line(mapping=aes(x=Year, y=lambda_fd, linetype=energy)) +
  scale_linetype_manual(values=c(2,3), name="") +
  facet_grid(Country ~ flavor) +
  scale_x_continuous(breaks=c(1960, 1980, 2000)) +
  xlab("") + 
  ylab(expression(lambda)) +
  xy_theme()
@
